---
title: "downsampling TCGA, WES-Other and PCAWG cohorts for validation and testing MESiCA"
author: "Adar Yaacov"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tibble)
library(ggplot2)
library(reshape2)
library(pROC)
library(precrec)
source("utils.R")
```

# downsample tcga
```{r}
wes_tcga_sig_label_mut <- read.table("../data/TCGA_mutation_classes_signature_annotations_filtered.tsv", sep = "\t", header = T, quote = "")
table(wes_tcga_sig_label_mut$sig)
length(unique(wes_tcga_sig_label_mut$sample))

#random sampling example of 1 mutation
wes_tcga_sig_label_mut_rS <- wes_tcga_sig_label_mut %>% 
  group_by(sample) %>% 
  slice_sample(n = 1)

```

#example - predict tcga random downsampled samples
```{r}
preds_tcga <- cos_pred_fun(sig_labels, 
                           unique(wes_tcga_sig_label_mut_rS$sample),
                           wes_tcga_sig_label_mut_rS,
                           creating_muts = TRUE, 
                           vep = FALSE)

preds_tcga_label <- left_join(preds_tcga[[1]], 
                              wes_tcga_sig_label_mut_rS[,c("sample","sig")] %>% group_by(sample) %>% summarise_all(max), 
                              by = "sample")
preds_binary_tcga_label <- left_join(preds_tcga[[3]], 
                              wes_tcga_sig_label_mut_rS[,c("sample","sig")] %>% group_by(sample) %>% summarise_all(max), 
                              by = "sample")
```

# collector - 20% percent downsampling
```{r}
# run 10 simulations of random 20% mutations and collect results

agg_aucs <- NULL

set.seed(1234) #for reproduceability 
for(i in 1:10){
  print(sprintf("Starting iteration %i", i))
  #random sampling
  wes_tcga_sig_label_mut_rS <- sig_wes_tcga_label_no45_no27_new_mut %>%
    group_by(sample) %>%
    slice_sample(prop = 0.2)
  
  #Finished sampling, starting predictions
  preds_tcga <- cos_pred_fun_efficient(sig_labels, 
                             unique(wes_tcga_sig_label_mut_rS$sample), wes_tcga_sig_label_mut_rS, 
                             creating_muts = TRUE, vep = FALSE, emb_len = 200)
  
  preds_tcga_label <- left_join(preds_tcga[[1]], 
                              wes_tcga_sig_label_mut_rS[,c("sample","sig")] %>% group_by(sample) %>%
                                summarise_all(max), 
                              by = "sample")
  
  preds_binary_tcga_label <- left_join(preds_tcga[[3]], 
                              wes_tcga_sig_label_mut_rS[,c("sample","sig")] %>% group_by(sample) %>%
                                summarise_all(max), 
                              by = "sample")

  #save to disk
  # write.csv(preds_tcga_label, sprintf("20pct_sampling/random_sampling_tcga_20pct_muts_results_%i.csv",i))
  # write.csv(preds_binary_tcga_label, sprintf("20pct_sampling/random_sampling_tcga_10_muts_results_binary_%i.csv",i))
  
  preds_binary_tcga_label_f <- confmat_preds(preds_binary_tcga_label)
  preds_binary_tcga_label_f <- preds_binary_tcga_label_f %>%
    filter(!sig %in% c("other")) %>%
    filter(!pred %in% c("other", "No prediction"))
  
  for(x in sig_labels[-6]){
    roc <- evalmod(labels = ifelse(preds_binary_tcga_label_f$sig == x, 1,0), 
                 scores = ifelse(preds_binary_tcga_label_f[,x] == "Yes", 1, 0))

    agg_aucs = rbind(agg_aucs,
                     cbind(sig = x,
                           AUROC = round(auc(roc)$aucs[1],digits = 3),
                           AUPRC = round(auc(roc)$aucs[2],digits = 3),
                           run = i))
  }
}
```

plot average AUC PRC 20%
```{r}
agg_aucs_mean <- as.data.frame(agg_aucs)
agg_aucs_mean$AUROC <- as.numeric(agg_aucs_mean$AUROC)
agg_aucs_mean$AUPRC <- as.numeric(agg_aucs_mean$AUPRC)

agg_aucs_mean <- agg_aucs_mean %>%
  group_by(sig) %>%
  summarize_all(mean)

gauc <- ggplot(agg_aucs_mean, aes(x=sig, y=AUROC)) + geom_col() + 
  scale_y_continuous(limits = c(0,1)) +
  theme_classic() +
  geom_text(aes(label = round(AUROC, digits = 3)), hjust = 1, angle=90, size=2.5, color="white") + 
  theme(text = element_text(size = 9, face = "bold"),
        legend.title = element_blank(),
        axis.title.y = element_text(size = 10),
        plot.margin = margin(0,0,0,0, "cm"),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, size=8))

gprc <- ggplot(agg_aucs_mean, aes(x=sig, y=AUPRC)) + geom_col() + 
  scale_y_continuous(limits = c(0,1)) +
  theme_classic() +
  geom_text(aes(label = round(AUPRC, digits = 3)), hjust = 1, angle=90, size=2.5, color="white") + 
  theme(text = element_text(size = 9, face = "bold"),
        legend.title = element_blank(),
        plot.margin = margin(0.2,0,0,0, "cm"),
        axis.title.y = element_text(size = 10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, size=8))

# tiff("figures/tcga_bench_DS_20pct_20rounds_w_label_180722_AUC_mean_2880_50rel.tiff", res = 300, width = 640, height = 960)
gridExtra::grid.arrange(gauc, gprc, ncol = 1)
# dev.off()
```

# Create high quality samples benchmark set - 2880 samples
```{r}
wes_tcga_sig_label_2k <- sig_wes_tcga_label[[2]] %>%
  left_join(sig_wes_tcga %>% 
              mutate(sum_mut = rowSums(sig_wes_tcga[,4:68])) %>%
              select(`Sample Names`, sum_mut),
            by = c('Sample Names'))
colnames(wes_tcga_sig_label_2k)[31] = "Tobacco"

wes_tcga_sig_label_2k_50 <- NULL
for(x in sig_labels[c(-3,-6)]){
  wes_tcga_sig_label_2k_50 <- rbind(
    wes_tcga_sig_label_2k_50, 
    wes_tcga_sig_label_2k[wes_tcga_sig_label_2k[,x] > 0.5,]
  )
}

wes_tcga_sig_label_2k_50 <- rbind(
  wes_tcga_sig_label_2k_50, 
  wes_tcga_sig_label_2k[wes_tcga_sig_label_2k[,"Clock_SBS5"] > 0.75,])
  
wes_tcga_sig_label_mut_50 <- wes_tcga_sig_label_mut %>%
  filter(sample %in% wes_tcga_sig_label_2k_50$`Sample Names`)

```

# collector - 1 to 15 mutations, benchmark set
```{r}
agg_scores <- NULL
for(n in 1:15){
  print(sprintf("Starting 20 times sampling of %i mutations", n))
  for(i in 1:10){
    #random sampling
    wes_tcga_sig_label_mut_rS <- wes_tcga_sig_label_mut_50 %>% 
      group_by(sample) %>%
      slice_sample(n = n)
  
    #Finished sampling, starting predictions
    preds_tcga_7 <- cos_pred_fun_efficient(sig_labels, unique(wes_tcga_sig_label_mut_rS$sample), wes_tcga_sig_label_mut_rS,
                             creating_muts = TRUE, vep = FALSE, emb_len = 200, sim_num = 0.6)

    preds_tcga_label_7 <- left_join(preds_tcga_7[[1]], 
                                wes_tcga_sig_label_mut_rS[,c("sample","sig")] %>% group_by(sample) %>%
                                  summarise_all(max), 
                                by = "sample")
    preds_binary_tcga_label_7 <- left_join(preds_tcga_7[[3]], 
                                wes_tcga_sig_label_mut_rS[,c("sample","sig")] %>% group_by(sample) %>%
                                  summarise_all(max), 
                                by = "sample")

    #save to disk
    # write.csv(preds_tcga_label_7, sprintf("1to15_sampling_2880/random_sampling_tcga_%i_muts_results_%i_040822.csv",n,i))
    # write.csv(preds_binary_tcga_label_7, sprintf("1to15_sampling_2880/random_sampling_tcga_%i_muts_results_binary_%i_040822.csv",n,i))
    
    preds_binary_tcga_label_f <- confmat_preds(preds_binary_tcga_label_7)
    preds_binary_tcga_label_f <- preds_binary_tcga_label_f %>%
      filter(!sig %in% c("other")) %>%
      filter(!pred %in% c("other", "No prediction"))
  
    conf <- as.data.frame(caret::confusionMatrix(
      as.factor(preds_binary_tcga_label_f$pred),
      as.factor(preds_binary_tcga_label_f$sig))$byClass)
  
    conf <- conf %>%
      dplyr::select(Sensitivity, Specificity, `Pos Pred Value`, `Neg Pred Value`) %>% 
      rownames_to_column("sig")
  
    agg_scores = rbind(agg_scores,
                       cbind(conf, run = i, mut_num = n))
  }
}
```

average 1 to 15 results
```{r}
agg_scores_mean <- agg_scores %>%
  group_by(sig, mut_num) %>%
  summarise_all(mean)
agg_scores_mean$sig <- gsub("Class: ", "", agg_scores_mean$sig)

# tiff("figures/ppv_1to15_grid_average_per_run_2880_50rel.tiff", res = 300, width = 1640, height = 1200)
ggplot(agg_scores_mean_50rel, 
       aes(x=mut_num, y=`Pos Pred Value`, group=1)) + 
  geom_line(color="#00A9FF",size=0.7) + 
  geom_point(size = 1.25) +
  facet_wrap(~sig, nrow = 2)+#, scales = "free") + 
  scale_x_continuous(breaks = c(seq(2,14,2))) +
  scale_y_continuous("Positive Predicted Value", 
                     breaks = seq(0,1,0.1), limits = c(0,1)) + 
  theme_bw() + xlab("Number of random mutations") + 
  theme(
    plot.margin = margin(0,1,0,0.15),
    panel.spacing = unit(0.1,"lines"),
    text = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14)
  ) 
# dev.off()
```

```{r}
sig_metrics <- melt(agg_scores_mean_50rel %>% 
                      filter(sig == "MMR") %>% 
                      select(!run), id.vars = c("mut_num","sig"))

# tiff("figures/MMR_1to15_grid_average_per_run_2880_50rel.tiff", res = 300, width = 1440, height = 640)
ggplot(sig_metrics, aes(x=mut_num, y=value)) + geom_line(color="orange",size=0.7) + geom_point(size = 1.25, alpha=0.75) + facet_grid(~variable) +
  scale_x_continuous(breaks = seq(1,15,2)) +
  scale_y_continuous("MMR signature", breaks = seq(0,1,0.1), limits = c(0,1)) +
  theme_bw() + xlab("Number of random mutations") + 
  theme(
    plot.margin = margin(0,0,0,0.05),
    panel.spacing = unit(0.2,"lines"),
    text = element_text(size = 9, face = "bold"),
    strip.text.x = element_text(size = 10),
    axis.title = element_text(size = 12)
  ) 
# dev.off()

```

#results analysis - model's "mistakes" in UV, MMR, POLE and APOBEC
POLE
```{r pole}
# pole <- agg_preds_clean %>% filter(sig == "POLE")
preds_20pct <- read.csv("20pct_sampling/random_sampling_tcga_10_muts_results_binary.csv", row.names = 1)
preds_20pct <- confmat_preds(preds_20pct)

pole <- preds_20pct %>% 
  filter(sig == "POLE")
pole <- pole %>% 
  left_join(sig_wes_tcga, by = c("sample"="Sample Names"))

pole$Cancer <- ifelse(pole$`Cancer Types` %in% c("Uterus-AdenoCa", "ColoRect-AdenoCa"), "Uterus/Colon", "Other")
pole$`MESiCA-POLE+` <- pole$POLE

g <- ggplot(pole, aes(x=reorder(sample,SBS10a+SBS10b), y=log2(SBS10a+SBS10b), col = `MESiCA-POLE+`, shape = Cancer))+
  geom_point(alpha = 0.7) + ylab("log2(SBS10a/b contribution)") + xlab("SBS10a/b samples by PCAWG") + 
  scale_y_continuous(breaks = seq(0,20,2)) + 
  theme_classic() + 
  theme(
    text = element_text(size = 7.5, face = "bold"),
    axis.text.x = element_blank(),
    legend.position = c(0.79,0.36),
    legend.spacing = unit(0.033,"cm"),
    legend.key.size = unit(0.25,"cm")
  )
  
# tiff("figures/pole_mistakes_20pct_average.tiff", res = 300, width = 580, height = 540)
g
# dev.off()
```

UV
```{r UV}
preds_20pct <- read.csv("20pct_sampling/random_sampling_tcga_10_muts_results_binary.csv", row.names = 1)

uv <- preds_20pct %>% 
  filter(sig == "UV")
uv <- uv %>% 
  left_join(sig_wes_tcga, by = c("sample"="Sample Names"))

uv$Cancer <- ifelse(uv$`Cancer Types` %in% c("Skin-Melanoma"), "Skin-Melanoma", "Other")
uv$`MESiCA-UV+` <- uv$UV

g <- ggplot(uv, aes(x=reorder(sample,SBS7a+SBS7b+SBS7c+SBS7d), y=log2(SBS7a+SBS7b+SBS7c+SBS7d), col = `MESiCA-UV+`, shape = Cancer))+
  geom_point(alpha = 0.7) + ylab("log2(UV-SBSs contribution)") + xlab("UV+ samples by PCAWG") + 
  scale_y_continuous(breaks = seq(0,20,2)) + 
  theme_classic() + 
  theme(
    text = element_text(size = 7.5, face = "bold"),
    axis.text.x = element_blank(),
    legend.position = c(0.79,0.25),
    legend.spacing = unit(0.033,"cm"),
    legend.key.size = unit(0.25,"cm")
  )
  
# tiff("figures/uv_misses_20pct_average.tiff", res = 300, width = 1440, height = 720)
g
# dev.off()

# tiff("figures/uv_misses_only_misses_20pct_average.tiff", res = 300, width = 720, height = 480)
ggplot(uv %>% filter(`MESiCA-UV+` != "Yes"), aes(x=reorder(sample,SBS7a+SBS7b+SBS7c+SBS7d), y=log2(SBS7a+SBS7b+SBS7c+SBS7d), col =Cancer, ))+
  geom_point(alpha = 0.7) + ylab("log2(UV-SBSs contribution)") + xlab("UV+ samples by PCAWG") + 
  scale_y_continuous(limits = c(0,9), breaks = seq(0,20,2)) + 
  theme_bw() + 
  theme(
    text = element_text(size = 7.5, face = "bold"),
    axis.text.x = element_blank(),
    legend.position = c(0.25,0.8),
    legend.spacing = unit(0.033,"cm"),
    legend.key.size = unit(0.25,"cm")
  )
# dev.off()

```

# PCAWG WGS
#downsaple pcawg benchmark set 
```{r}
sig_wgs_pcawg_list <- process_sig(sig_wgs_pcawg, ngs = "wgs")

sig_wgs_pcawg_label <- sig_wgs_pcawg_list[[2]]
colnames(sig_wgs_pcawg_label)[31] <- "Tobacco"

sig_wgs_pcawg_label_50 <- NULL
for(x in sig_labels[c(-3,-6)]){
  sig_wgs_pcawg_label_50 <- rbind(
    sig_wgs_pcawg_label_50, 
    sig_wgs_pcawg_label[sig_wgs_pcawg_label[,x] > 0.5,]
  )
}

sig_wgs_pcawg_label_50 <- rbind(
  sig_wgs_pcawg_label_50, 
  sig_wgs_pcawg_label[sig_wgs_pcawg_label[,"Clock_SBS5"] > 0.75,])

pcawg_sig_label_mut_50 <- pcawg_sig_label_mut %>%
  filter(sample %in% sig_wgs_pcawg_label_50$`Sample Names`) %>%
  filter(!sig %in% c("MMR","POLE","Clock_SBS1","HRD"))
```

# downsample pcawg
```{r}
agg_scores_pcawg_50rel <- NULL

pcawg_sig_label_mut_f <- pcawg_sig_label_mut %>%
  filter(sample %in% sig_wgs_pcawg_label_50$`Sample Names`)

pcawg_sig_label_mut_f1 <- pcawg_sig_label_mut_f[pcawg_sig_label_mut_f$sample %in% pcawg_sig_label_mut_freq_f$Var1,]
pcawg_sig_label_mut_f2 <- pcawg_sig_label_mut_f[!pcawg_sig_label_mut_f$sample %in% pcawg_sig_label_mut_freq_f$Var1,]


set.seed(1234)
for(i in 1:10){
  print(sprintf("Starting iteration %i", i))
  #random sampling
  pcawg_sig_label_mut_rS_1 <- pcawg_sig_label_mut_f1 %>%
    group_by(sample) %>%
    slice_sample(prop = 0.002)
  
  pcawg_sig_label_mut_rS_2 <- pcawg_sig_label_mut_f2 %>%
    group_by(sample) %>%
    slice_sample(n = 10)

  pcawg_sig_label_mut_rS <- rbind(pcawg_sig_label_mut_rS_1, pcawg_sig_label_mut_rS_2)

  print(sprintf("Finished sampling, starting predictions, iteration %i", i))
  preds_pcawg <- cos_pred_fun(sig_labels, unique(
    pcawg_sig_label_mut_rS$sample), pcawg_sig_label_mut_rS,
                           creating_muts = TRUE, vep = FALSE, emb_len = 200)
  preds_pcawg_label <- left_join(preds_pcawg[[1]], 
                              pcawg_sig_label_mut_rS[,c("sample","sig")] %>% group_by(sample) %>%
                                summarise_all(max), 
                              by = "sample")
  preds_binary_pcawg_label <- left_join(preds_pcawg[[3]], 
                              pcawg_sig_label_mut_rS[,c("sample","sig")] %>% group_by(sample) %>%
                                summarise_all(max), 
                              by = "sample")
  
  #save to disk
  # write.csv(preds_pcawg_label, sprintf("pcawg_rand_sampling_10/random_sampling_pcawg_%i_muts_results_%i_50rel.csv",10,i))
  # write.csv(preds_binary_pcawg_label, sprintf("pcawg_rand_sampling_10/random_sampling_pcawg_%i_muts_results_binary_%i_50rel.csv",10,i))

  preds_binary_pcawg_label_f <- confmat_preds(preds_binary_pcawg_label)
  preds_binary_pcawg_label_f <- preds_binary_pcawg_label_f %>%
    filter(!sig %in% c("other", "HRD")) %>%
    filter(!pred %in% c("other", "No prediction","HRD"))

  conf <- as.data.frame(caret::confusionMatrix(
    as.factor(preds_binary_pcawg_label_f$pred),
    as.factor(preds_binary_pcawg_label_f$sig))$byClass)

  conf <- conf %>%
    dplyr::select(Sensitivity, Specificity, `Pos Pred Value`, `Neg Pred Value`) %>% 
    rownames_to_column("sig")

  agg_scores_pcawg_50rel = rbind(agg_scores_pcawg_50rel,
                     cbind(conf, run = i, mut_num = n))
}

```

#mean metrics
```{r}

agg_scores_pcawg_002_or_10_50rel <- agg_scores_pcawg_50rel %>% 
  group_by(sig, mut_num) %>%
  summarise_all(mean)
agg_scores_pcawg_002_or_10_50rel$sig <- gsub("Class: ", "", agg_scores_pcawg_002_or_10_50rel$sig)

```

#pred pcawg w/o sampling
```{r}
preds_pcawg_by_tcga <- cos_pred_fun(sig_labels, unique(pcawg_sig_label_mut$sample),
                                    pcawg_sig_label_mut, creating_muts = T,
                                    vep = F, emb_len = ncol(embeddings))
```


# pred WES-OTHER cohorts BENCHMARK
#collector - 1 to 15
```{r}
agg_scores_wes_other <- NULL

set.seed(12345)
for(n in 1:15){
  print(sprintf("Starting 20 times sampling of %i mutations", n))
  for(i in 1:10){
    #random sampling
    wes_other_sig_label_mut_rS <- sig_wes_other_label_mut_bench %>% 
      group_by(sample) %>%
      slice_sample(n = n)
  
    #Finished sampling, starting predictions
    preds_other_7 <- cos_pred_fun_efficient(sig_labels, unique(wes_other_sig_label_mut_rS$sample), wes_other_sig_label_mut_rS,
                             creating_muts = TRUE, vep = FALSE, emb_len = 200, sim_num = 0.6)
    preds_other_label_7 <- left_join(preds_other_7[[1]], 
                                wes_other_sig_label_mut_rS[,c("sample","sig")] %>% group_by(sample) %>%
                                  summarise_all(max), 
                                by = "sample")
    preds_binary_other_label_7 <- left_join(preds_other_7[[3]], 
                                wes_other_sig_label_mut_rS[,c("sample","sig")] %>% group_by(sample) %>%
                                  summarise_all(max), 
                                by = "sample")

    #save to disk
    # write.csv(preds_other_label_7, sprintf("1to15_sampling_wes_other_bench/random_sampling_wes_other_%i_muts_results_%i.csv",n,i))
    # write.csv(preds_binary_other_label_7, sprintf("1to15_sampling_wes_other_bench/random_sampling_wes_other_%i_muts_results_binary_%i.csv",n,i))

    preds_binary_other_label_f <- confmat_preds(preds_binary_other_label_7)
    preds_binary_other_label_f <- preds_binary_other_label_f %>%
      filter(!sig %in% c("other")) %>%
      filter(!pred %in% c("other", "No prediction"))
  
    conf <- as.data.frame(caret::confusionMatrix(
      as.factor(preds_binary_other_label_f$pred),
      as.factor(preds_binary_other_label_f$sig))$byClass)
  
    conf <- conf %>%
      dplyr::select(Sensitivity, Specificity, `Pos Pred Value`, `Neg Pred Value`) %>% 
      rownames_to_column("sig")
  
    agg_scores_wes_other = rbind(agg_scores_wes_other,
                       cbind(conf, run = i, mut_num = n))
  }
}
```

#mean
```{r}

agg_scores_mean_tcga_bench <- mean_per_experiment(agg_scores)
# write.csv(agg_scores_wes_other_bench_new_mean, "agg_scores_1to15_TCGA_bench_MEAN_050822.csv")

agg_scores_wes_other_bench_new_mean <- mean_per_experiment(agg_scores_wes_other)
# write.csv(agg_scores_wes_other_bench_new_mean, "agg_scores_1to15_wes_other_bench_MEAN_050822.csv")
```

plot mean
```{r}
# tiff("figures/ppv_1to15_grid_average_per_run_TCGA_WES_other_bench.tiff", res = 300, width = 1640, height = 1200)
ggplot(agg_scores_wes_other_bench_new_mean, 
       aes(x=mut_num, y=mean_ppv, group=1)) + geom_line(color="#00A9FF",size=0.7) +  geom_point(size = 1.25) +
  facet_wrap(~sig, nrow = 2)+
  scale_x_continuous(breaks = c(seq(2,14,2))) +
  scale_y_continuous("Positive Predicted Value", 
                     breaks = seq(0,1,0.1), limits = c(0,1)) + 
  theme_bw() + xlab("Number of random mutations") + 
  theme(
    plot.margin = margin(0,1,0,0.15),
    panel.spacing = unit(0.1,"lines"),
    text = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14)
  ) +
  geom_errorbar(aes(ymin=mean_ppv-ci_ppv, ymax=mean_ppv+ci_ppv), width=.1) + 
  geom_line(data = agg_scores_mean_tcga_bench, aes(x=mut_num, y=mean_ppv, group=1), color="red",size=0.7) + 
  geom_point(data = agg_scores_mean_tcga_bench, aes(x=mut_num, y=mean_ppv, group=1), size = 1.25) + 
  geom_errorbar(data = agg_scores_mean_tcga_bench, aes(ymin=mean_ppv-ci_ppv, ymax=mean_ppv+ci_ppv), width=.1)
# dev.off()

# tiff("figures/sens_1to15_grid_average_per_run_TCGA_WES_other_bench.tiff", res = 300, width = 1640, height = 1200)
ggplot(agg_scores_wes_other_bench_new_mean, 
       aes(x=mut_num, y=mean_sens, group=1)) + geom_line(color="#00A9FF",size=0.7) +  geom_point(size = 1.25) +
  facet_wrap(~sig, nrow = 2)+#, scales = "free") + 
  scale_x_continuous(breaks = c(seq(2,14,2))) +
  scale_y_continuous("Sensitivity", 
                     breaks = seq(0,1,0.1), limits = c(0,1)) + 
  theme_bw() + xlab("Number of random mutations") + 
  theme(
    plot.margin = margin(0,1,0,0.15),
    panel.spacing = unit(0.1,"lines"),
    text = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14)
  ) +
  geom_errorbar(aes(ymin=mean_sens-ci_sens, ymax=mean_sens+ci_sens), width=.1) + 
  geom_line(data = agg_scores_mean_tcga_bench, aes(x=mut_num, y=mean_sens, group=1), color="red",size=0.7) + 
  geom_point(data = agg_scores_mean_tcga_bench, aes(x=mut_num, y=mean_sens, group=1), size = 1.25) + 
  geom_errorbar(data = agg_scores_mean_tcga_bench, aes(ymin=mean_sens-ci_sens, ymax=mean_sens+ci_sens), width=.1)

# dev.off()

# tiff("figures/npv_1to15_grid_average_per_run_TCGA_WES_other_bench.tiff", res = 300, width = 1640, height = 1200)
ggplot(agg_scores_wes_other_bench_new_mean, 
       aes(x=mut_num, y=mean_npv, group=1)) + geom_line(color="#00A9FF",size=0.7) +  geom_point(size = 1.25) +
  facet_wrap(~sig, nrow = 2)+#, scales = "free") + 
  scale_x_continuous(breaks = c(seq(2,14,2))) +
  scale_y_continuous("Negative Predicted Value", 
                     breaks = seq(0,1,0.1), limits = c(0,1)) + 
  theme_bw() + xlab("Number of random mutations") + 
  theme(
    plot.margin = margin(0,1,0,0.15),
    panel.spacing = unit(0.1,"lines"),
    text = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14)
  ) +
  geom_errorbar(aes(ymin=mean_npv-ci_npv, ymax=mean_npv+ci_npv), width=.1) + 
  geom_line(data = agg_scores_mean_tcga_bench, aes(x=mut_num, y=mean_npv, group=1), color="red",size=0.7) + 
  geom_point(data = agg_scores_mean_tcga_bench, aes(x=mut_num, y=mean_npv, group=1), size = 1.25) + 
  geom_errorbar(data = agg_scores_mean_tcga_bench, aes(ymin=mean_npv-ci_npv, ymax=mean_npv+ci_npv), width=.1)

# dev.off()

# tiff("figures/spec_1to15_grid_average_per_run_TCGA_WES_other_bench.tiff", res = 300, width = 1640, height = 1200)
ggplot(agg_scores_wes_other_bench_new_mean, 
       aes(x=mut_num, y=mean_spec, group=1)) + geom_line(color="#00A9FF",size=0.7) +  geom_point(size = 1.25) +
  facet_wrap(~sig, nrow = 2)+#, scales = "free") + 
  scale_x_continuous(breaks = c(seq(2,14,2))) +
  scale_y_continuous("Specificity", 
                     breaks = seq(0,1,0.1), limits = c(0,1)) + 
  theme_bw() + xlab("Number of random mutations") + 
  theme(
    plot.margin = margin(0,1,0,0.15),
    panel.spacing = unit(0.1,"lines"),
    text = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14)
  ) +
  geom_errorbar(aes(ymin=mean_spec-ci_spec, ymax=mean_spec+ci_spec), width=.1) + 
  geom_line(data = agg_scores_mean_tcga_bench, aes(x=mut_num, y=mean_spec, group=1), color="red",size=0.7) + 
  geom_point(data = agg_scores_mean_tcga_bench, aes(x=mut_num, y=mean_spec, group=1), size = 1.25) + 
  geom_errorbar(data = agg_scores_mean_tcga_bench, aes(ymin=mean_spec-ci_spec, ymax=mean_spec+ci_spec), width=.1)

# dev.off()
```

plot mean apobec mmr
```{r}

sens_labels <- c("Sensitivity", "Specificity", "NPV")
names(sens_labels) <- c("mean_sens", "mean_spec", "mean_npv")

# tiff("figures/apobec_mmr_metrics_1to15_TCGA_WES_OTHER_bench.tiff", res = 300, width = 980, height = 1640)
ggplot(
  melt(agg_scores_wes_other_bench_new_mean[agg_scores_wes_other_bench_new_mean$sig %in% c("APOBEC","MMR"),c(1:3, 11, 15)], 
       id.vars = 1:2), 
  aes(x=mut_num, y=value, group=1)) + geom_line(color="#00A9FF",size=0.7) +  geom_point(size = 1.25) +
  facet_grid(variable~sig, 
             labeller = labeller(variable = sens_labels),
             switch="y") + 
           theme_bw() + xlab("Number of random mutations") + 
           theme(
               plot.margin = margin(0,1,0,0.15),
               panel.spacing = unit(0.1,"lines"),
               text = element_text(size = 12, face = "bold"),
               axis.title = element_text(size = 14),
               strip.placement = "outside",
               strip.text.x = element_text(size = 12),
               strip.text.y = element_text(size = 15),
               strip.background.y = element_rect(colour = "white", fill="white")
           ) +   scale_x_continuous(breaks = c(seq(2,14,2))) +
    scale_y_continuous(NULL, 
                       breaks = seq(0,1,0.1), limits = c(0,1)) +
  geom_line(data = melt(agg_scores_mean_tcga_bench[agg_scores_mean_tcga_bench$sig %in% c("APOBEC","MMR"),c(1:3, 11, 15)], id.vars = 1:2), 
            aes(x=mut_num, y=value, group=1), color="red",size=0.7) + 
  geom_point(data = melt(agg_scores_mean_tcga_bench[agg_scores_mean_tcga_bench$sig %in% c("APOBEC","MMR"),c(1:3, 11, 15)], id.vars = 1:2),
             aes(x=mut_num, y=value, group=1), size = 1.25)

# dev.off()
```

# collector WES Other - 20% percent benchmark set
```{r}
wes_other_agg_aucs_bench <- NULL

set.seed(1234)
for(i in 1:10){
  #randoom sampling
  wes_other_sig_label_mut_rS <- sig_wes_other_label_mut_bench %>% 
    group_by(sample) %>%
    slice_sample(prop = 0.2)

  preds <- cos_pred_fun_efficient(sig_labels, 
                             unique(wes_other_sig_label_mut_rS$sample), wes_other_sig_label_mut_rS, 
                             creating_muts = TRUE, vep = FALSE, emb_len = 200)
  
  preds_label <- left_join(preds[[1]], 
                              wes_other_sig_label_mut_rS[,c("sample","sig")] %>% group_by(sample) %>%
                                summarise_all(max), 
                              by = "sample")
  
  preds_binary_label <- left_join(preds[[3]], 
                              wes_other_sig_label_mut_rS[,c("sample","sig")] %>% group_by(sample) %>%
                                summarise_all(max), 
                              by = "sample")

  #save to disk
  # write.csv(preds_label, sprintf("20pct_sampling_wes_other/random_sampling_wes_other_20pct_bench_%i.csv",i))
  # write.csv(preds_binary_label, sprintf("20pct_sampling_wes_other/random_sampling_wes_other_20pct_binary_bench_%i.csv",i))
  
  preds_binary_label_f <- confmat_preds(preds_binary_label)
  preds_binary_label_f <- preds_binary_label_f %>%
    filter(!sig %in% c("other")) %>%
    filter(!pred %in% c("other", "No prediction"))
  
  for(x in sig_labels[-6]){
    roc <- evalmod(labels = ifelse(preds_binary_label_f$sig == x, 1,0), 
                 scores = ifelse(preds_binary_label_f[,x] == "Yes", 1, 0))

    wes_other_agg_aucs_bench = rbind(wes_other_agg_aucs_bench,
                     cbind(sig = x,
                           AUROC = round(auc(roc)$aucs[1],digits = 3),
                           AUPRC = round(auc(roc)$aucs[2],digits = 3),
                           run = i))
  }
}
```

plot average AUC PRC 20%
```{r}
wes_other_agg_aucs_bench_mean <- as.data.frame(wes_other_agg_aucs_bench)
wes_other_agg_aucs_bench_mean$AUROC <- as.numeric(wes_other_agg_aucs_bench_mean$AUROC)
wes_other_agg_aucs_bench_mean$AUPRC <- as.numeric(wes_other_agg_aucs_bench_mean$AUPRC)

wes_other_agg_aucs_bench_mean <- wes_other_agg_aucs_bench_mean %>%
  group_by(sig) %>%
  summarize_all(mean)

gauc <- ggplot(wes_other_agg_aucs_bench_mean, aes(x=sig, y=AUROC)) + geom_col() + 
  scale_y_continuous(limits = c(0,1)) +
  theme_classic() +
  geom_text(aes(label = round(AUROC, digits = 3)), hjust = 1, angle=90, size=2.5, color="white") + 
  theme(text = element_text(size = 9, face = "bold"),
        legend.title = element_blank(),
        axis.title.y = element_text(size = 10),
        plot.margin = margin(0,0,0,0, "cm"),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, size=8))

gprc <- ggplot(wes_other_agg_aucs_bench_mean, aes(x=sig, y=AUPRC)) + geom_col() + 
  scale_y_continuous(limits = c(0,1)) +
  theme_classic() +
  geom_text(aes(label = round(AUPRC, digits = 3)), hjust = 1, angle=90, size=2.5, color="white") + 
  theme(text = element_text(size = 9, face = "bold"),
        legend.title = element_blank(),
        plot.margin = margin(0.2,0,0,0, "cm"),
        axis.title.y = element_text(size = 10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, size=8))

# tiff("figures/wes_other_agg_aucs_prcs_AUC_bench_050822.tiff", res = 300, width = 640, height = 960)
gridExtra::grid.arrange(gauc, gprc, ncol = 1)
# dev.off()
```
